### 2018-10-18 1/exams/m1.tex
Midterm 1 source.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 10 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 10 \\ \hline
  7 & 20 \\ \hline \hline
  Total: & 90\\
  \hline
\end{tabular}
\end{center}

@@

\begin{verbatim}
(define R 9)
(define (w z) (= R z))
(cond [(w 2) 4]
      [else 9])
\end{verbatim}

@@

(cond [(w 2) 4] [else 9])     -->[plug]
       ^^^^^
(cond [(= R 2) 4] [else 9])   -->[const]
          ^
(cond [(= 9 2) 4] [else 9])   -->[arith]
       ^^^^^^^
(cond [#false 4] [else 9])    -->[cond]
^^^^^^^^^^^^^^^^^^^^^^^^^^
(cond [else 9])               -->[cond]
^^^^^^^^^^^^^^^
9
\end{SaveVerbatim}

@@

\begin{verbatim}
(define-struct run (time up score))
\end{verbatim}

@@

- make-run  : Constructor
- run-time  : Accessor
- run-up    : Accessor
- run-score : Accessor
- run?      : Predicate
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 3 (10 points).}
%
The graphical text editor you developed in a recent assignment has
attracted the interest of investors.  But before they write you a
check, they want to make sure you can handle fancy operations for
expert text editor users.  They ask if you can implement a \emph{swap}
operation, which will swap content to the left of the cursor with the
content to the right..  To satisfy your potential investors, you
define the following function for swapping strings at a given
position:
\begin{verbatim}
;; swap : String Integer -> String
;; Swap string to the left of i with string to the right.
;; Assume (<= 0 i (string-length s)).
(check-expect (swap "" 0) "")
(check-expect (swap "ab" 1) "ba")
(check-expect (swap "bannana" 3) "nanaban")
(define (swap s i) ...)
\end{verbatim}

\noindent
Give a correct definition for \verb|swap|.  (You only need to
provide code, not the design steps.)

@@

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Use the following data definition for representing names:
\begin{verbatim}
;; A Name is a (make-name String String)
;; Interp: a person's full (first and last) name.
;; Both strings must contain at least one letter.
(define-struct name (first last))
\end{verbatim}

\noindent
Write a program, \verb|name-width|, that determines the width needed
to print someone's name on a letter using a mono-typed font (meaning
each letter has the same width).  The program can calculate this width
based on the length of the person's first and last name, plus the
width of a space to go in between.  Assume each letter is 10 pixels
wide.  For example, the name Bob Harper requires 100 pixels; the name
Bo Jack requires 70 pixels.  You may start from the following header:

\begin{verbatim}
;; name-width : Name -> Number
;; Computes width needed to print name in mono-type font
\end{verbatim}

(check-expect (name-width (make-name "Bob" "Harper")) 100)
(check-expect (name-width (make-name "Bo" "Jack")) 70)
(define (name-width n)
  (* 10 (+ 1 (string-length (name-first n))
             (string-length (name-last n)))))
\end{SaveVerbatim}

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of non-negative numbers in the list.  For example, if the list
contains 3, 4, -1, 0, -2, and 5, the count of non-negative numbers is
4.  You may use the following data definition:

@@

\noindent
{\bf Problem 6 (10 points).} You've been hired by the US Department of
Transportation to design traffic simulation software, which includes
simulating traffic lights.  Discovering the simple design from 131A is
insufficient, you come up with the following representation, which
makes it possible to model not just the color, but also how much
longer the current light has before changing:

\begin{verbatim}
;; A Light is one of:
;; - (make-red Natural)
;; - (make-yellow Natural)
;; - (make-green Natural)
;; Interp: color and remaining duration (in ticks) of a light 
(define-struct red (count))
(define-struct yellow (count))
(define-struct green (count))
\end{verbatim}
(Recall that a \verb|Natural| is a natural number, i.e. a non-negative
integer.)

@@

\noindent
{\bf Problem 6(a).} Write a template for \verb|Light| functions.

@@

(define (light-temp l)
  (cond [(red? l) (... (red-count l) ...)]
        [(yellow? l) (... (yellow-count l) ...)]
        [(green? l) (... (green-count l) ...)]

\vskip 15em

@@

\noindent
{\bf Problem 6(b).} Write a stub for the following function:

\begin{verbatim}
;; tock : Light -> Light
;; Advance the light by one tick of time
(check-expect (tock (make-red 10)) (make-red 9))
(check-expect (tock (make-red 0)) (make-green 10))
\end{verbatim}

@@

(define (tock l) (make-red 0))

@@

\noindent
{\bf Problem 7 (20 points).}
%
Over the summer, you take an internship at the University of
Maryland's Center for Bioinformatics and Computational Biology (CBCB).
Your first task deals with computing information about \emph{DNA
  strands}.  A DNA strand is an arbitrarily long sequence of
\emph{nucleotides}, which are either: C, G, A, or T.  Design a program
that, given a DNA strand, computes the number of times the G
nucleotide occurs in it.  For example, if the strand is AGCTTTGA, your
program should produce 2.

@@

;; A Nucleotide is one of:
;; - "C"
;; - "G"
;; - "A"
;; - "T"
;; Interp: DNA nucleotide C,G,A,T.

;; A DNA is one of:
;; - '()
;; - (cons Nucleotide DNA)
;; Interp: a DNA strand of nucleotides

;; count-g : DNA -> Natural
;; Count the number of occurrence of G in given DNA strand
(check-expect (count-g '()) 0)
(check-expect (count-g (cons "A" (cons "G" (cons "G" '())))) 2)
(define (count-g dna)
  (cond [(empty? dna) 0]
        [(cons? dna)
         (+ (if (string=? (first dna) "G") 1 0)
            (count-g (rest dna)))]))

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2018 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 40 \\ \hline
  Total: & 90\\
  \hline
\end{tabular}
\end{center}

@@

\begin{verbatim}
(define Q 1)
(define (w z) (< (string-length z) 20))
(w (cond [(= 1 Q) "fred"]
         [else 9]))
\end{verbatim}

@@

(w (cond [(= 1 Q) "fred"] [else 9])) -->[const]
               ^
(w (cond [(= 1 1) "fred"] [else 9])) -->[arith]
          ^^^^^^^
(w (cond [#true "fred"] [else 9]))   -->[cond]
   ^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^
(w "fred")                           -->[plug]
^^^^^^^^^^
(< (string-length "fred") 20))       -->[arith]
   ^^^^^^^^^^^^^^^^^^^^^^
(< 5 20)                             -->[arith]
#true
\end{SaveVerbatim}

@@

\begin{verbatim}
(define-struct dr (who strange))
\end{verbatim}

@@

- make-dr    : Constructor
- dr-who     : Accessor
- dr-strange : Accessor
- dr?        : Predicate
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 3 (15 points).}
%
You've been hired by the CMNS development office and been given their
existing software for spamming potential donors.  It contains the following
data definition:
\begin{verbatim}
;; A Person is a (make-name Title String String)
;; Interp: a person's title and first and last names.
(define-struct name (title first last))

;; A Title is one of:
;; - "r"      Interp: Mr.
;; - "s"              Ms.
;; - "d"              Dr.
\end{verbatim}

\noindent
Finish the design of the following function for creating letter
openings:
\begin{verbatim}
;; dear : Person -> String
;; Create a letter opening for the given person.
(check-expect (dear (make-name "d" "Minnie" "Maisy"))
              "Dear Dr. Maisy:")
(check-expect (dear (make-name "r" "Fred" "Rogers"))
              "Dear Mr. Rogers:")
(check-expect (dear (make-name "s" "Miriam" "Maisel"))
              "Dear Ms. Maisel:")
\end{verbatim}

@@

\newpage
\noindent
{\bf Problem 4 (15 points).}  
%
Design a program that takes a list of numbers and produces the sum of
positive numbers in the list.  For example, if the list contains 3, 4,
-1, 0, -2, and 5, the sum of positive numbers (3+4+5) is 12.  You may
use the following data definition:

@@

\noindent
{\bf Problem 5 (40 points).} You've been hired by Nintendo to design
their next blockbuster game: Dots.  It consists of, well, dots: lots
and lots of dots, each of which occupies a coordinate in space.  You
come up with the following data definitions for representing a single
dot and a collection of dots:

\begin{verbatim}
;; A Dots is one of:
;; - empty
;; - (cons Dot Dots)
;; Interp: collection of an arbitrary number of dots

;; A Dot is a (make-posn Number Number)
;; Interp: a dot's position in space
\end{verbatim}

@@

\noindent
{\bf Problem 5(a) [10 points].} Write templates for \verb|Dot| and
\verb|Dots| functions.

@@

(define (dot-template d)
  (... (posn-x d)
       (posn-y d)
       ...))

(define (dots-template ds)
  (cond [(empty? ds) ...]
        [(cons? ds) (... (dot-template (first ds))
                         (dots-template (rest ds)) ...)]))
\end{SaveVerbatim}

\vskip 18em

@@

\noindent
{\bf Problem 5(b) [10 points].} Write a stub for the following function:

\begin{verbatim}
;; nw : Dots -> Dots
;; Move given dots in the northwest direction one unit
(check-expect (nw empty) empty)
(check-expect (nw (cons (make-posn 3 4) empty))
              (cons (make-posn 4 5) empty))
\end{verbatim}

@@

(define (nw ds) empty)

@@

\noindent
{\bf Problem 5(c) [20 points].}
%
Design two functions: {\tt dot-flip} reflects a given dot over the
y-axis, for example, $(3,4)$ becomes $(-3,4)$, and {\tt dots-flip}
reflects a collection of dots over the y-axis.
\begin{SaveVerbatim}{VerbEnv}

@@

;; dot-flip : Dot -> Dot
;; Flip dot over y-axis
(check-expect (dot-flip (make-posn 3 4)) (make-posn -3 4))
(define (dot-flip d)
  (make-posn (- (posn-x d)) (posn-y d)))

;; dots-flip : Dots -> Dots
;; Flip dots over y-axis
(check-expect (dots-flip empty) empty)
(check-expect (dots-flip (cons (make-posn 3 4) empty))
              (cons (make-posn -3 4) empty))
(define (dots-flip ds)
  (cond [(empty? ds) empty]
        [(cons? ds) 
         (cons (dot-flip (first ds))
               (dots-flip (rest ds)))]))                   
\end{SaveVerbatim}
